---
title: "Mixed Models"
output:
  word_document: default
  pdf_document: default
date: "2022-10-17"
---
--------------------------------------------------------------------------------
Mixed Effects Models
--------------------------------------------------------------------------------
Need these packages:

install.packages("remotes")
remotes::install_github("easystats/report")

To use the the "report" library.

```{r, warning=FALSE}
x <- c("ggmap", "rgdal", "rgeos", "maptools", "dplyr", "tidyr", "tmap", "sf", "MASS", "rgeos", "gmt", "lme4", "lmerTest", "ggplot2", "sf", "tidyverse", "gganimate", "report")
lapply(x, library, character.only = TRUE) 
```

```{r}
# Laura's PC
# dataPop = read.csv("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/populationLong.csv")

# Laura's laptop example below
# dataPop = read.csv("C:\\Users\\Laura\\OneDrive - Western Sydney University\\Spring 2022\\Final Project\\populationLong.csv")

# My's laptop Example Below
# dataPop = read.csv("C:/Users/viquy/Documents/GitHub/Maths_Project/scripts/datasets/populationLong.csv")

dataPop = read.csv("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/populationLong.csv")
```


This is simply to make the data frame easier to read. We're not using these columns so why waste space.
```{r}
dataPop = subset(dataPop, select = -c(numChange,perChange) )
```

Create density column and calculate density for each SA2 each year. 
Create yearsince2000 column.
```{r}
dataPop$density <- 0.0    
dataPop$density <- dataPop$pop/dataPop$area
dataPop$yearsince2000 <- dataPop$year-2000
```

```{r}
dataPop$trans_density <- sqrt(dataPop$density)
```

```{r}
# Laura's PC
# syd <- st_read("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/SA2_2016_AUST.shp")

# Laura's laptop example below
# syd <- st_read("C:\\Users\\Laura\\OneDrive - Western Sydney University\\Spring 2022\\SA2_2016_AUST.shp")

# My's laptop Example Below
# syd <- st_read("C:/Users/viquy/Documents/GitHub/Maths_Project/scripts/datasets/SA2_2016_AUST.shp")

syd <- st_read("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/SA2_2016_AUST.shp")
greaterSyd <- syd[syd$GCC_NAME16=="Greater Sydney",]
```

```{r}
centr <- st_centroid(greaterSyd)
centroid_lat <- unlist(lapply(centr$geometry,function(l) l[2]))
centroid_long <- unlist(lapply(centr$geometry,function(l) l[1]))
```

================================================================================
Here is where we define our centroids. 
Started with Sydney and Parramatta. 
The distances are calculated for each SA2 and the normalized via a log 
tansformation.
================================================================================

```{r}
greaterSyd$sydist <- geodist(-33.8688,151.2093,centroid_lat,centroid_long)
greaterSyd$parradist <- geodist(-33.8148,151.0017,centroid_lat,centroid_long)
dataPop$sydist <- greaterSyd$sydist[match(dataPop$name,greaterSyd$SA2_NAME16)]
dataPop$parradist <- greaterSyd$parradist[match(dataPop$name,greaterSyd$SA2_NAME16)]
dataPop$logsydist <- log(greaterSyd$sydist[match(dataPop$name,greaterSyd$SA2_NAME16)])
dataPop$logparradist <- log(greaterSyd$parradist[match(dataPop$name,greaterSyd$SA2_NAME16)])
```



```{r}
# Drop SA2's without syddist (Only 1 sample)
dataPop <- dataPop[!is.na(dataPop$logsydist),]
length(unique(dataPop$logsydist))
length(dataPop$logsydist)
```

So all together we have 3588 samples.

================================================================================

Mixed-Models:
A mixture of fixed and random. 

Fixed effects model.
```{r}
fit_fixed <- lm(trans_density~yearsince2000*logparradist+yearsince2000*logsydist,data=dataPop)
summary(fit_fixed)
#report(fit_fixed)
```

================================================================================
Output Summary Interpretation:
I'm assuming we can interpret this in the same way we approached the linear
model. The only difference being the ways in which we examine the 
variables like yearsince2000:parradist and yearsince2000:sydist.

If so we can say this model performs poorly compared to the multiple linear 
regression and linear models above.

^ This performs similarly to the previous models. Accounting for roughly
59.51% of the variance in the data
The only concerning thing is that in both models yearsince2000:parradist
and yearsince2000:sydist are not apparently significant given their p-values. 



This model doesn't allow us to answer the proposed question so consider 
rejecting it. The MLR proposed in the previous RMD performs better and would 
be used instead. 
================================================================================



================================================================================
Mixed Effect Model, Random Intercepts


Random intercepts, in our model, are defined when we seperate the data into 
levels, 1 for each SA2 name. These intercepts all have similar, if not identical 
slopes, however their intercepts are all unique. 
================================================================================

```{r}
fit_randint <- lmer(trans_density~yearsince2000*logparradist*logsydist+(1|name),data=dataPop)
summary(fit_randint)
#report(fit_randint)
```


================================================================================
Summary analysis: 
Refer to report() for a more detail description of the model. 
This model faces similar issues to our previous one in that it does not
specifically answer the question. 
We need to consider the years 2020 and 2021 to truly see their impact on 
density.
================================================================================


================================================================================
Mixed Effect Model, Random Slopes


Random slopes, in our model, are defined when we separate the data into 
levels, over yearsince2000 for each SA2 name.
================================================================================
```{r}
fit_randslope1 <- lmer(trans_density~yearsince2000*parradist*sydist+(yearsince2000|name),data=dataPop)
summary(fit_randslope1)
# report(fit_randslope1)
```



================================================================================
Comparing The Models. 
We can compare the two using an ANOVA. 
================================================================================

```{r}
anova(fit_randint,fit_randslope1)
```

Random slope model significantly better fit.


================================================================================
Let's create logical variables for the years 2020 and 2021. This states, if the
sample is from 2020, the column, year2020 = TRUE. If not, it's FALSE. 
year2021 is created in the same way.
================================================================================
```{r}
dataPop$year2020 = (dataPop$yearsince2000 == 20)
dataPop$year2021= (dataPop$yearsince2000 == 21)
```

Check this worked.
```{r}
head(dataPop,30)
```

================================================================================
Random Slopes Model 
This model only considers Sydney distance. This choice was made after using a 
similar interaction model that also included Parramatta. 
================================================================================
```{r}
slopes_years_syd = lmer(trans_density ~ logsydist*yearsince2000 + logsydist*year2020 +  logsydist*year2021 + (yearsince2000|name),data=dataPop)
summary(slopes_years_syd)
#report(slopes_years_syd)
```


================================================================================
We see is this model that all the variables are highly significant. 
Also note the variance is what we'd expect from the data. It's very high 
when comparing different SA2 regions but the variance within the regions and 
from year to year is much smaller. 

This is why we choose a mixed model, it accounts from the variences among the
features.
================================================================================





================================================================================
Prediction With Random Slopes Model

Create a new df that holds samples to be plotted based on the model.
Here we have a data point (Surry Hills) close the Sydney, 2kms, and one far 
away from Sydney (Avoca Beach), 50kms.
================================================================================

```{r}
# Find specific distance SA2 regions from Sydney 
dataPop[order(dataPop$logsydist, decreasing = FALSE), ]
```
Darlinghurst, Surry Hills, Lidcombe ,Avoca Beach, Wentworth Falls
```{r}
df = expand.grid(logsydist=c(0.3208779, 0.6340370,2.7107842,3.876380,4.360940), yearsince2000=1.0:21.0)
df
```


Now use to ifelse statements to establish if the sample if from 2020 or 2021
```{r}
df$year2020 = ifelse(df$yearsince2000 == 20, TRUE, FALSE)
df$year2021 = ifelse(df$yearsince2000 == 21, TRUE, FALSE)
```


```{r}
df_predict = predict(slopes_years_syd, df, re.form=~0)
```


```{r}
df_predict**2
```

col= ifelse(df_predict**2 >= 1000, "#56B4E9","#CC99FF")
```{r}
plot(df$yearsince2000, df_predict**2, col = c("#56B4E9","#4169E1","#D2691E" ,"#CC99FF", "#6A5ACD"), xlab="Years Since 2000", ylab="Density Predictions", pch=16)
```


```{r}
slope_syd = coef(slopes_years_syd)$name
#slope_syd
```


```{r}
greaterSyd$slope_syd <- slope_syd$`(Intercept)`[match(unique(greaterSyd$SA2_NAME16),row.names(slope_syd))]
qtm(shp = greaterSyd, fill = c("slope_syd"), fill.palette = "Purples", title = "Years Since 2000 Coefficients")
```




```{r}
ints_years_syd = lmer(trans_density ~ logsydist*yearsince2000 + logsydist*year2020 +  logsydist*year2021 + (1|name),data=dataPop)
summary(ints_years_syd)
#report(slopes_years_syd)
```

```{r}
anova( slopes_years_syd,ints_years_syd)
```


================================================================================
Try the same slope model as above with Parramatta as our main point of reference. 
================================================================================

```{r}
slopes_years_parra = lmer(trans_density ~ logparradist*yearsince2000 + logparradist*year2020 +  logparradist*year2021 + (yearsince2000|name),data=dataPop)
summary(slopes_years_parra)
#report(slopes_years_parra)
```

This model also performs quite well. The main difference is the interaction 
effect between logparradist and year2020TRUE
Let's try a similar prediction as done with the last model.


```{r}
# Find specific distance SA2 regions from Parramatta. 
dataPop[order(dataPop$logparradist, decreasing = FALSE), ]
```


```{r}
# Sample 1 is North Parramatta, which is 1.993615kms away from the Parramatta centroid. 
# Sample 2 is Katoomba - Leura, which is 65.16629kms away.
df_parra = expand.grid(logparradist=c(0.6899496, 0.8246439, 2.3740355, 4.517657, 4.176942), yearsince2000=1.0:21.0)

df_parra$year2020 = ifelse(df_parra$yearsince2000 == 20, TRUE, FALSE)
df_parra$year2021 = ifelse(df_parra$yearsince2000 == 21, TRUE, FALSE)
```

```{r}
df_predict_parra = predict(slopes_years_parra, df_parra, re.form=~0)

plot(df_parra$yearsince2000, df_predict_parra**2)
```

Frick yeah. 
This seems to be a consistent pattern. The closer to Parramatta an SA2 is, the 
more its density drops off in 2020 and 2021.


```{r}

```




================================================================================
OLD MODELS/OLD PLOTS
Keep for reference, but know that we're not using them. 
================================================================================

Random slope with year= 2020 and year = 2021 as predictors
```{r}
slopes_years = lmer(trans_density ~ logsydist*logparradist*yearsince2000 + logsydist*logparradist*year2020 +  logsydist*logparradist*year2021 + (yearsince2000|name),data=dataPop)
#summary(slopes_years)
slope_coef = coef(slopes_years)$name
#random_coef
#row.names(random_coef)
```


This is showing the initial intercepts. Need to re do!
```{r}
greaterSyd$coef_slope <- slope_coef$`(Intercept)`[match(unique(greaterSyd$SA2_NAME16),row.names(slope_coef))]
qtm(shp = greaterSyd, fill = c("coef_slope"), fill.palette = "Purples", title = "Coeficient Heat Map (Slopes)")
```

================================================================================
Intercepts - Sydney only model.

================================================================================

```{r}
intercepts_years_syd = lmer(trans_density ~ logsydist*yearsince2000 + logsydist*year2020 +  logsydist*year2021 + (1|name),data=dataPop)
summary(intercepts_years_syd)
```


================================================================================

```{r}
anova(intercepts_years_syd, slopes_years_syd)
```

================================================================================



================================================================================
Visualization
- Plots I'm trying to emulate from github.
================================================================================


Need this chunk to work for the next chunk
```{r}
fixed_effects = fixef(slopes_years_parra)
random_effects = 
  slope_coef %>%
  mutate(name = factor(unique(dataPop$name)))
```


Since this SA2 data is from 2016 I can only really use SA2's with names consistent from 2016.
Drop missing ones?
This is an issue with trying to reproduce the next plot as well


Maybe try this with a simpler mixed model first

```{r}
dataPop %>% 
  ggplot(aes(logsydist, trans_density)) +
  geom_point(aes(color = 1), alpha = .25) +
  geom_abline(
    aes(
      intercept = fixed_effects['(Intercept)'],
      slope = fixed_effects['Sydney Dist']),
    color = 'darkred',
    size = 2) +
  geom_abline(
    aes(
      intercept = `(Intercept)`, 
      slope = logsydist,
      color = 1),
    size = .5,
    alpha = .25,
    data = random_effects) 
```



```{r}
# This is how I want to plot the random slopes
# Pretty sure
dataPop %>% 
  ggplot(aes(logsydist*logparradist*yearsince2000 + logsydist*logparradist*year2020 +  logsydist*logparradist*year2021, trans_density)) +
  geom_point(aes(color = name), alpha = .25) +
  geom_abline(
    aes(
      intercept = fixed_effects['(Intercept)'],
      slope = fixed_effects['Name']),
    color = 'darkred',
    size = 2) +
  geom_abline(
    aes(
      intercept = '(Intercept)', 
      slope = logsydist*logparradist*yearsince2000 + logsydist*logparradist*year2020 +  logsydist*logparradist*year2021,
      color = name),
    size = .5,
    alpha = .25,
    data = dataPop) 
```

================================================================================
Sample code

model_fit = lmer(trans_density ~ logsydist*logparradist*yearsince2000 + logsydist*logparradist*year2020 +  logsydist*logparradist*year2021 + (1|name),data=dataPop)
model_lm = lm(trans_density ~ logsydist + yearsince2000)

```{r}
model_lm = lm(trans_density ~ logsydist + yearsince2000, data = dataPop)
```


```{r}
gpa_lm_by_group0 = lmList(trans_density ~ yearsince2000 | name, dataPop)
gpa_lm_by_group  = coef(gpa_lm_by_group0)
gint = 
  data_frame(Mixed=coef(model_fit)$name[,1], Separate=gpa_lm_by_group[,1]) %>% 
  gather(key=Model, value=Intercept) %>% 
  ggplot(aes(x=Intercept)) +
  geom_density(aes(color=Model, fill=Model), alpha=.25) +
  scale_color_viridis_d(begin = .25, end = .75) +
  scale_fill_viridis_d(begin = .25, end = .75) +
  ggtitle('Intercepts') +
  labs(x='', y='') +
  xlim(c(1.5,4)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.key.size=unit(2, 'mm'),
    legend.title=element_text(size=8),
    legend.text=element_text(size=8),
    legend.box.spacing=unit(0, 'in'),
    legend.position=c(.85,.75)
    )
gslopes = 
  data_frame(Mixed=coef(gpa_mixed)$student[,2], Separate=gpa_lm_by_group[,2]) %>% 
  gather(key=Model, value=Occasion) %>% 
  ggplot(aes(x=Occasion)) +
  geom_density(aes(color=Model, fill=Model), alpha=.25, show.legend=F) +
  scale_color_viridis_d(begin = .25, end = .75) +
  scale_fill_viridis_d(begin = .25, end = .75) +
  ggtitle('Slopes for occasion') +
  labs(x='', y='') +
  xlim(c(-.2,.4)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
library(patchwork)
gint + gslopes
```



```{r}
greaterSyd
```

```{r}
qtm(shp = greaterSyd, fill = c("parradist"), fill.palette = "Blues")
```

================================================================================
Sample code

model_fit = lmer(trans_density ~ logsydist*logparradist*yearsince2000 + logsydist*logparradist*year2020 +  logsydist*logparradist*year2021 + (1|name),data=dataPop)
model_lm = lm(trans_density ~ logsydist + yearsince2000)

```{r}
model_lm = lm(trans_density ~ logsydist + yearsince2000, data = dataPop)
```


```{r}
gpa_lm_by_group0 = lmList(trans_density ~ yearsince2000 | name, dataPop)
gpa_lm_by_group  = coef(gpa_lm_by_group0)

gint = 
  data_frame(Mixed=coef(model_fit)$name[,1], Separate=gpa_lm_by_group[,1]) %>% 
  gather(key=Model, value=Intercept) %>% 
  ggplot(aes(x=Intercept)) +
  geom_density(aes(color=Model, fill=Model), alpha=.25) +
  scale_color_viridis_d(begin = .25, end = .75) +
  scale_fill_viridis_d(begin = .25, end = .75) +
  ggtitle('Intercepts') +
  labs(x='', y='') +
  xlim(c(1.5,4)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.key.size=unit(2, 'mm'),
    legend.title=element_text(size=8),
    legend.text=element_text(size=8),
    legend.box.spacing=unit(0, 'in'),
    legend.position=c(.85,.75)
    )

gslopes = 
  data_frame(Mixed=coef(gpa_mixed)$student[,2], Separate=gpa_lm_by_group[,2]) %>% 
  gather(key=Model, value=Occasion) %>% 
  ggplot(aes(x=Occasion)) +
  geom_density(aes(color=Model, fill=Model), alpha=.25, show.legend=F) +
  scale_color_viridis_d(begin = .25, end = .75) +
  scale_fill_viridis_d(begin = .25, end = .75) +
  ggtitle('Slopes for occasion') +
  labs(x='', y='') +
  xlim(c(-.2,.4)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


library(patchwork)
gint + gslopes
```




