---
title: "Population Data Manipulation"
author: "Laura Pierce"
date: "2022-09-09"
---

This is how I manipulated the population data set into long format.

Libraries needed
```{r}
library("geosphere")
library("ggplot2")
library("ggmap")
library("tidyr")
library("ggplot2")
library('MASS')
library("sf")
library("rgeos")
library("gmt")
library("lme4")
library("lmerTest")
```

Read in ABS population data from 2001-2021
```{r}
df = read.csv("C:\\Users/Laura/OneDrive - Western Sydney University\\Spring 2022\\Mathematics Project\\Data Sets (Need to organise)\\population_mod.csv")
head(df,5)
```

For loop to correct the year column names. Want them to be numeric and not include 'X'
```{r}
for (x in 4:24){
  name = 1997 + x
  colnames(df)[x] <- name
}
head(df,5)
```

Convert to long format first before dropping 0 population values. Then I can simply do it by the resulting population column.


```{r}
df = df %>% pivot_longer(cols=c('2001', '2002', '2003', '2004', '2005', '2006', '2007', '2008', '2009', '2010',  '2011', '2012', '2013', '2014', '2015', '2016', '2017', '2018', '2019', '2020', '2021'),
                    names_to='year',
                    values_to='Pop')


head(df,5)
dim(df)
```


Drop columns where population = 0 to make further calculations easier.
```{r}
data = df[!grepl(0, df$Pop),]
dim(data)
head(data,5)
```
Drop columns that aren't needed
```{r}
data = subset(data, select = -c(X,SA2_Code,Population_Density) )
```

```{r}
head(data,5)
```

Save this data frame as long data set.
```{r}
write.csv(data,"C:\\Users/Laura/OneDrive - Western Sydney University\\Spring 2022\\Mathematics Project\\Data Sets (Need to organise)\\populationLong.csv", row.names = FALSE)
```




