---
title: "Project Data Visualisation"
output: pdf_document
date: "2022-09-19"
---

library(tidyr)
library(MASS)
library(sf)
library(rgeos)
library(gmt)
library(lme4)
library(lmerTest)
library(ggplot2)

Libaries
```{r}
x <- c("ggmap", "rgdal", "rgeos", "maptools", "dplyr", "tidyr", "tmap", "sf", "MASS", "rgeos", "gmt", "lme4", "lmerTest", "ggplot2")
lapply(x, library, character.only = TRUE) 
```

```{r}
df = read.csv("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/populationLong.csv")
df$density <- 0.0    
df$density <- df$pop/df$area
df$yearsince2000 <- df$year-2000
head(df,5)
```



Load in shape files and population data
```{r}
syd <- st_read("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/SA2_2016_AUST.shp")
```




```{r}
greaterSyd <- syd[syd$GCC_NAME16=="Greater Sydney",]
plot(greaterSyd)
```

Compare density with something else owo
qtm(shp = lnd, fill = c("Partic_Per", "Pop_2001"), fill.palette = "Blues", ncol = 2)



```{r}
qtm(shp = greaterSyd, fill = c("AREASQKM16"), fill.palette = "Blues")
```



```{r}
syd <- syd[syd$GCC_NAME16=="Greater Sydney",]
centr <- st_centroid(syd)
```


```{r}
centroid_lat <- unlist(lapply(centr$geometry,function(l) l[2]))
centroid_long <- unlist(lapply(centr$geometry,function(l) l[1]))
```


```{r}
syd$sydist <- geodist(-33.8688,151.2093,centroid_lat,centroid_long)
syd$parradist <- geodist(-33.8148,151.0017,centroid_lat,centroid_long)
df$sydist <- syd$sydist[match(df$name,syd$SA2_NAME16)]
df$parradist <- syd$parradist[match(df$name,syd$SA2_NAME16)]

df$logsydist <- log(syd$sydist[match(df$name,syd$SA2_NAME16)])
df$logparradist <- log(syd$parradist[match(df$name,syd$SA2_NAME16)])

```



```{r}
qtm(shp = syd, fill = c("sydist"), fill.palette = "Blues", title = "NSW Distance Map")
```


```{r}
syd_30 <- syd[syd$sydist < 30,]
qtm(shp = syd_30, fill = c("sydist", "AREASQKM16"), fill.palette = "Purples", ncol = 2, title = c('Distance', 'Area'))
```




```{r}
syd_20 <- syd[syd$sydist < 20,]
qtm(shp = syd_20, fill = c("sydist", "AREASQKM16"), fill.palette = "Blues", ncol = 2, title = c('Distance', 'Area'))
```

How to subset based on multiple condition

```{r}
pop2002 <- df[(df$year == 2016) | (df$year == 2017),]
head(pop2002,40)
```


```{r}
syd$popNum <- pop2002$pop[match(syd$SA2_NAME16, pop2002$name)]
```



```{r}
pop2002[pop2002$name=="Wyong",]
```


```{r}
qtm(shp = syd, fill = c("popNum"), fill.palette = "Purples") +
      tm_legend(legend.position = c("left", "top"),
          main.title = "2016 Population in Greater Sydney",
          main.title.position = "middle"
          )
```


?qtm
```{r}
RColorBrewer::display.brewer.all()
```


Fix this up and try new stuff
```{r}
pop2002 <- df[df$year == 2016,]
head(pop2002,40)
```


```{r}
syd$popNum <- pop2002$pop[match(syd$SA2_NAME16, pop2002$name)]
```


Goals: 
Could create for loop for plotting separate years. 
Look into faceting 

Load in original Russell data
```{r}
df2 = read.csv("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/Russell Data.csv")
head(df2)
```


```{r}
syd2 <- st_read("C:\\Users/Owner/OneDrive - Western Sydney University/Spring 2022/Mathematics Project/Data Sets (Need to organise)/SA2_2016_AUST.shp")
```



syd$popNum <- pop2002$pop[match(syd$SA2_NAME16, pop2002$name)]

```{r}
syd2$years <- df2$TIME_FY[match(syd2$SA2_NAME16, df2$Region)]
```

```{r}
syd2$migration <- df2$Value[match(syd2$SA2_NAME16, df2$Region)]
```



```{r}
head(syd2$AREASQKM16)
```




```{r}
tm_shape(syd2) +
    tm_polygons("AREASQKM16", title="Uhhh") +
    tm_facets("years")
```


```{r}
qtm(shp = syd2, fill = c("migration"), fill.palette = "Blues", title = "NSW migration Map")
```
